\section{Core Engine Features}

This section describes a few less technical aspects of the game engine. For
example, the Entities section (section \ref{sec:entities}), describes of a
modular organization of a small stats engine, that may be defined dynamically by
the user.

\subsection{Entities}
\label{sec:entities}

We will now discuss about a possible organization for the data that an entity
mahy hold and manipulate. Essentially an entity should be composed of `hard
facts', and `modular facts'. Hard facts of an entity would be things such as the
entitie's \textit{strength}, \textit{magic}, and other similar information. Such
statistical information should be represented as real numbers: $\mathbb{R}$.
When it is time for implementation, you might wish to prefer \texttt{float}
types instead of \texttt{double}, as the extended precision of the latter is not
required. Each of these attributes, come bundled with a label, which will be
used to tell the player what kind of attribute the player is observing. We can
therefore express this as seen in the algebraic specifications
\ref{eq:attribute}, and \ref{eq:entity}.

\begin{equation}
\begin{split}
  \label{eq:attribute}
  Label\colon String \\
  Attribute\colon Label \times \mathbb{R}
\end{split}
\end{equation}

Now let us describe the module-plugin organization. Since we want to exhibit
different behaviors in the stats of an Entity, we want the Entity to maintain
a list of these behaviours. Notice that these behaviors take in a type
\texttt{Entity}, and map it to an \texttt{Entity} as well. So these behaviors
should essentially alter the state of the \texttt{Entity}, when fired up. So for
instance, you could create different functions for a \textit{Fighter}, and
define the attributes affected on a level up. This does not exclude the
possibility of adding more functions, and composing a function out of different
behaviours. For example, if the player is cursed, there could be a function that
gives a penalty per level up. We show such an organization in
Specifications~\ref{eq:funcmodules}, and later how it is bound to the entity in
\ref{eq:entity}.

\begin{equation}
\begin{split}
  \label{eq:funcmodules}
  f \colon Entity \to Entity \\
  MODULES_{seq}\colon \langle f_1, f_2, ..., f_n \rangle
\end{split}
\end{equation}

Another essential aspect to RPGs is the equipment system, which allows the
player to tweak the performance of the \texttt{Entity}. This is simply a type,
which holds a label, and any number of \texttt{Attribute}s. This way we can
specify modifiers for things like strength, intelligence, and more, as well as
add other things like elemental damage. We can see this in specification
\ref{eq:equipment}.

\begin{equation}
\begin{split}
  \label{eq:equipment}
  Equipment\colon Label \times \langle Attribute \rangle
\end{split}
\end{equation}

Now, for the sake of flexibility, we should provide a set of functions that will
perform checks each time a piece of equipment is attempted to be equiped for the
character. This should take in an \texttt{Entity}, an \texttt{Equipment} and
yield a \texttt{Boolean}, as demonstrated in
specification~\ref{eq:equipmentcheck}. This way, we define a sequence which
contains all these constraints as well, with the name
\texttt{EquipmentConstraints}. For setups where each character can equip
anything, then this may be set as an empty list.

\begin{equation}
\begin{split}
  \label{eq:equipmentcheck}
  f\colon Entity \times Equipment \to Boolean \\
  EquipConstraints_{seq} = \langle f_1, f_2, ..., f_n \rangle
\end{split}
\end{equation}

We should also specify some container in the implementation, to contain one item
from the enumeration shown in specification~\ref{eq:entityclass}. An interesting
question is if it would be possible to mix any number of classes, to get a mixed
class, and in turn automatically constrain skills or buffs on an entity. For now
we stick to single class assignment as seen in
specification~\ref{eq:entityclass}. And finally we can define our composite type
\texttt{Entity} with the specification shown in \ref{eq:entity}.

\begin{equation}
  \label{eq:entityclass}
  EntityClass = \{ Fighter, Mage, Thief, Assassin, Archer, ... \} \\
\end{equation}

\begin{equation}
\begin{split}
  \label{eq:entity}
  Entity\colon Label \times
    EntityClass \times
    \langle Attribute \rangle \times
    \langle Equipment \rangle \\
    \times MODULE_{seq} \times EquipmentConstraints_{seq}
\end{split}
\end{equation}

Note that since $MODULE_{seq}$, and $EquipmentConstraints_{seq}$ are both
sequences of functions, with same type, which implies that we can compose them,
in order to get a wanted result. So for instance, the final form of the entity,
could be represented as shown in equation~\ref{eq:entitycomposition}. In
equation \ref{eq:equipconstrcomp} we can see a similar approach, to check that
any constraints that the entity may have, are satisfied or not, yielding a
boolean value in the end.

\begin{equation}
\begin{split}
  \label{eq:entitycomposition}
  Entity_{final} = (f_1 \circ f_2 \circ ... \circ f_n)(E_{simple}) \\
  where, \langle f_1, f_2, ..., f_n \rangle \in MODULE_{seq}
\end{split}
\end{equation}

\begin{equation}
\begin{split}
  \label{eq:equipconstrcomp}
  CanEquip = (f_1 \circ f_2 \circ ... \circ f_n)(Entity_{given}, Equipment_{try}) \\
  where, \langle f_1, f_2, ..., f_n \rangle \in EquipmentConstraints_{seq}
\end{split}
\end{equation}

\subsubsection{Examples}

We list a few examples, in this subsection, so that we animate the
specifications above. Let's say we have a fighter. We can create an entity as
shown in \ref{eq:createent}. Translating this to source code at this point would
be trivial, since each part so far has been well defined.

\begin{equation}
\begin{split}
  \label{eq:createent}
  fighter = Entity('Arnold', Fighter, \langle str, int, dex \rangle \\
    \langle sword, shield, boots \rangle, \langle f_{strbuff}, f_{intdebuff}
  \rangle, \langle f_{eqonlyfighter} \rangle)
\end{split}
\end{equation}

\subsection{Leveling System}

You can bind possibly a leveling system to the \texttt{Entity} type. It works
very similar to the buffs as we previously saw (binding functions to the entity,
and letting itself deal with itself). You could do this by binding the leveling
functions directly to the \texttt{Entity} data type, or making a data type to
wrap around the \texttt{Entity} data type. To keep things a little simple, it
would be easier to define a type \texttt{LevelingSystem}, and binding any
level-up strategies in there. So in essence, from section~\ref{sec:entities}, we
extend the specification, finally to the given specification
at~\ref{eq:entitylevelup}.

\begin{equation}
  \begin{split}
  \label{eq:entitylevelup}
  Entity\colon Label \times
    EntityClass \times
    \langle Attribute \rangle \times
    \langle Equipment \rangle \\
    \times MODULE_{seq} \times EquipmentConstraints_{seq} \\
    \times LevelingEngine
  \end{split}
\end{equation}

The \texttt{LevelingEngine} should aggregate functions whose type should be
$f\colon Entity \to Entity$. What this means is that these functions should
change the state of \texttt{Entity} whenever there is a levelup that happens.
The functions, for example that could happen for such an entity are things like
$increaseStr$, or $increaseDef$, or even $increaseAgility$. Now the way you
implement this depends on how you define your entities, and how you define your
attributes that your entity aggregates.

Implementation aside, the reader should notice that since we can aggregate
different functions this way, we could modularly define attribute increases per
level in the editor. For example for the $Fighter$ class we could aggregate a
$increaseStr$ function with special parameters to add extra points to strength
per level up. Or we could even aggregate the function a few times many to
achieve the same effect.

\subsection{Battle System}

In this section we describe how the Entity structure (section
\ref{sec:entities}), may be used in order to form a battle system.

\subsection{Storage of Assets}

In this section we describe possible strategies on how to persist the generated
assets by the game editor.
