\section{Glossary and Core Concepts}

We explain the meaning of some of the technical jargon that may be found in this
document, for the unfamiliar reader.

\begin{description}

\item {Map}: A part of the world the game takes place in. A map will contain
  information about what tilesets it's using, and the specific tiles used from
  that tileset. So in other words, each time a tile is set in the map, there
  should be a tuple stored somewhere, which contains an id of the tileset being
  used (or possibly a path), and an id which references the tile inside of that
  tileset.

\item {Map Layers}: A map can have many layers, and transparency of images must
  be supported. There are two reasons: you might want to have layers that add
  extra niceties in the form of graphics. For example seeing things from an
  overhead view, and having leaves of trees (or branches) be over a player, yet
  see the player under the branches. Another motivation would be that we can add
  obstructions on different layers, meaning that the player could go on
  different floors of a smaller building for example, while everything is in
  fact on only one map. For example think of a bridge, where the player can move
  under, and over.

\item {Tile}: a (usually) square graphic that can be used as a pattern to fill
  in the space of the map. For example a ``grass pattern'' can be expressed in
  one square graphic, and reused in order to build the map. Tiles exist on
  \textit{map layers}.

\item {Tileset}: A collection of tiles (as described above). Tilesets usually
  have a particular style, or respect a scene (castle tilesets, should have
  things that are found in castles, overworld should have grass and trees, etc).

%% TODO is this really the name I want?
\item {Entity}: An entity in a game, is anyting that might have some stats, and
  more information attached to it (usually these are monsters, can be nps, and
  the heroes the player can control).

\item {NPC}: Stands for `non-playable-character', and they are the people the
  main character finds in villages, or eslewhere, that usually offer helpful
  hints and tips.

\item {Pawn}: A pawn is a placeable `thing' in a map, on a layer. These differ
  to tiles, in the sense that they are animated, and more than simple tiles on a
  map. For example NPCs, treasure chests, and anything else that might be deemed
  special. Pawns should contain some scripted behavior, that can tie the game
  together. For example it should be possible to perform map transitions via
  some minor scripting language, via a Pawn. These functionalities should
  include more things like changing the sprite, and more.

\item {Portal}: A connecting unit between map `A' and map `B'. There might be
  many ways going about this --- maybe this is a special feature of a map, or
  there can be reusable functionality coded inside the engine, which would be
  used via a placed entity on the map, to perform this transition.


\end{description}

